
```{r}
source("config.R")
knitr::opts_chunk$set(dev = "png", dpi = 300, fig.path = file.path(DIR_WGCNA, "figures/"))
```

# CHUNK 1: Libraries
```{r setup, message = FALSE, warning = FALSE}
library(WGCNA)
library(edgeR)
library(limma)
library(clusterProfiler)
library(org.Hs.eg.db)
library(dplyr)
library(ggplot2)
library(reshape2)
library(tidyr)
library(ggrepel)
library(msigdbr)
library(fgsea)
library(patchwork)
library(scales)
```


# CHUNK 2: Paths
```{r paths}
COUNTDIR <- file.path(BASEDIR, "counts")
QCDIR    <- file.path(BASEDIR, "qc")
OUTDIR <- DIR_WGCNA
if (!dir.exists(OUTDIR)) dir.create(OUTDIR, recursive = TRUE)
cat("Output will be written to:", OUTDIR, "\n")
```

# CHUNK 3: Load raw data
```{r load-data}
raw_counts <- read.delim(file.path(COUNTDIR, "DEA_counts.tsv"),
                         row.names = 1, check.names = FALSE)
anno  <- read.delim(file.path(COUNTDIR, "gene_symbols.tsv"), check.names = FALSE)
meta0 <- read.delim(file.path(QCDIR, "DEA_samples_QCmetrics.tsv"),
                    check.names = FALSE)
cat("Raw counts:", nrow(raw_counts), "genes ×", ncol(raw_counts), "samples\n")
cat("Annotation:", nrow(anno), "genes\n")
cat("Metadata  :", nrow(meta0), "samples\n")
```


# CHUNK 4: Filter metadata
```{r filter-metadata}
meta <- meta0 %>%
  filter(case_control %in% c("case", "control"),
         tissue_pathology == "primary DCIS",
         is.na(final_qc) | trimws(as.character(final_qc)) == "") %>%
  distinct(sample_id, .keep_all = TRUE)
cat("Samples after filtering:", nrow(meta), "\n")
cat("  Case   :", sum(meta$case_control == "case"),    "\n")
cat("  Control:", sum(meta$case_control == "control"), "\n")
cat("\nBatch × case_control table:\n")
print(table(meta$batch, meta$case_control))
cat("\nFisher's exact test (avoids the chi-square zero-cell warning):\n")
print(fisher.test(table(meta$batch, meta$case_control)))
```


# CHUNK 5: Align counts to metadata
```{r align-samples}
shared_samples <- intersect(colnames(raw_counts), meta$sample_id)
counts <- raw_counts[, shared_samples]
meta   <- meta[match(shared_samples, meta$sample_id), ]
rownames(meta) <- meta$sample_id
stopifnot(identical(colnames(counts), meta$sample_id))
cat("Aligned matrix:", nrow(counts), "genes ×", ncol(counts), "samples\n")
```


# CHUNK 6: Biotype filter (protein_coding + lncRNA)
```{r biotype-filter}
keep_biotype <- anno$ensembl_gene_id[anno$gene_biotype %in% c("protein_coding", "lncRNA")]
counts <- counts[rownames(counts) %in% keep_biotype, ]
cat("After biotype filter:", nrow(counts), "genes\n")
```


# CHUNK 7: Expression filter (edgeR::filterByExpr)
```{r filter-expression}
group_factor <- factor(meta$case_control)
batch_factor <- factor(meta$batch)
y <- DGEList(counts = counts, group = group_factor)
keep <- filterByExpr(y, group = group_factor)
y <- y[keep, , keep.lib.sizes = FALSE]
cat("After expression filter:", nrow(y), "genes\n")
```


# CHUNK 8: TMM + voom 
```{r normalise-voom, fig.width = 7, fig.height = 5}
while (!is.null(dev.list())) dev.off()       # clear any stale devices
par(mfrow = c(1, 1), mar = c(5, 4, 4, 2) + 0.1)
y <- calcNormFactors(y, method = "TMM")
design <- model.matrix(~ 0 + group_factor)
colnames(design) <- levels(group_factor)
v <- voom(y, design, plot = TRUE)
cat("Voom matrix:", nrow(v$E), "genes ×", ncol(v$E), "samples\n")
```


# CHUNK 8A: PCA before batch correction
```{r pca-before, fig.width = 7, fig.height = 5}
pca_before <- prcomp(t(v$E), center = TRUE, scale. = FALSE)
var_before <- summary(pca_before)$importance[2, 1:5]
df_before  <- data.frame(PC1 = pca_before$x[, 1],
                         PC2 = pca_before$x[, 2],
                         batch = batch_factor,
                         case_control = group_factor)
print(ggplot(df_before, aes(PC1, PC2, colour = batch, shape = case_control)) +
        geom_point(size = 3) +
        labs(title = "PCA before batch correction",
             x = sprintf("PC1 (%.1f%%)", 100 * var_before[1]),
             y = sprintf("PC2 (%.1f%%)", 100 * var_before[2])) +
        theme_bw())
cat("BEFORE — PC1 ~ batch F =",
    summary(aov(pca_before$x[,1] ~ batch_factor))[[1]]$"F value"[1],
    "| PC1 ~ case F =",
    summary(aov(pca_before$x[,1] ~ group_factor))[[1]]$"F value"[1], "\n")
```


# CHUNK 8B: Batch correction
```{r batch-correct}
design_keep <- model.matrix(~ group_factor)
v$E <- removeBatchEffect(v$E, batch = batch_factor, design = design_keep)
cat("limma::removeBatchEffect applied on v$E.\n")
```


# CHUNK 8C: PCA after batch correction
```{r pca-after, fig.width = 7, fig.height = 5}
pca_after <- prcomp(t(v$E), center = TRUE, scale. = FALSE)
var_after <- summary(pca_after)$importance[2, 1:5]
df_after  <- data.frame(PC1 = pca_after$x[, 1],
                        PC2 = pca_after$x[, 2],
                        batch = batch_factor,
                        case_control = group_factor)
print(ggplot(df_after, aes(PC1, PC2, colour = batch, shape = case_control)) +
        geom_point(size = 3) +
        labs(title = "PCA after batch correction",
             x = sprintf("PC1 (%.1f%%)", 100 * var_after[1]),
             y = sprintf("PC2 (%.1f%%)", 100 * var_after[2])) +
        theme_bw())
cat("AFTER  — PC1 ~ batch F =",
    summary(aov(pca_after$x[,1] ~ batch_factor))[[1]]$"F value"[1],
    "| PC1 ~ case F =",
    summary(aov(pca_after$x[,1] ~ group_factor))[[1]]$"F value"[1], "\n")
```


# CHUNK 9: log2(RPKM+1)
```{r rpkm-biotype}
gene_lengths_all <- setNames(anno$width, anno$ensembl_gene_id)
gene_lengths_y   <- gene_lengths_all[rownames(y$counts)]
rpkm_matrix      <- rpkm(y, gene.length = gene_lengths_y)
log2rpkm_mat     <- log2(rpkm_matrix + 1)
log2rpkm_data    <- t(log2rpkm_mat)
```


# CHUNK 10: Transpose for WGCNA
```{r wgcna-input}
expression.data <- t(v$E)
cat("WGCNA input: samples =", nrow(expression.data),
    ", genes =", ncol(expression.data), "\n")
```

# CHUNK 11: goodSamplesGenes
```{r quality-check}
gsg <- goodSamplesGenes(expression.data, verbose = 3)
if (!gsg$allOK) {
  expression.data <- expression.data[gsg$goodSamples, gsg$goodGenes]
  meta <- meta[meta$sample_id %in% rownames(expression.data), ]
}
cat("Final matrix:", nrow(expression.data), "samples ×",
    ncol(expression.data), "genes\n")
```


# CHUNK 12: Sample clustering
```{r sample-clustering, fig.width = 14, fig.height = 5}
sampleTree <- hclust(dist(expression.data), method = "average")
par(mar = c(0, 4, 2, 0))
plot(sampleTree, main = "Sample clustering after limma batch correction",
     sub = "", xlab = "")
```


# CHUNK 13: Trait dataframe
```{r trait-setup}
to_na <- function(x, na_vals = c("", "NA", "N/A", "999", "unknown", "Unknown")) {
  x <- trimws(as.character(x)); x[x %in% na_vals] <- NA; x
}
recode_bin <- function(x, one_vals, zero_vals) {
  x <- to_na(x); out <- rep(NA_real_, length(x))
  out[tolower(x) %in% tolower(one_vals)]  <- 1
  out[tolower(x) %in% tolower(zero_vals)] <- 0
  out
}
recode_ord <- function(x, allowed) {
  x <- suppressWarnings(as.numeric(to_na(x))); x[!x %in% allowed] <- NA; x
}
allTraits <- data.frame(
  sample_id    = meta$sample_id,
  case_control = recode_bin(meta$case_control, "case", "control"),
  ER           = recode_ord(meta$er,  c(0, 1)),
  PR           = recode_ord(meta$pr,  c(0, 1)),
  HER2         = recode_ord(meta$her2, c(0, 1, 2)),
  radiotherapy = recode_ord(meta$radiotherapy, c(0, 1)),
  grade        = recode_ord(meta$grade, c(1, 2, 3)))
allTraits$HER2 <- ifelse(allTraits$HER2 == 2, NA_real_, allTraits$HER2)
Samples   <- rownames(expression.data)
traitRows <- match(Samples, allTraits$sample_id)
stopifnot(!any(is.na(traitRows)))
datTraits <- allTraits[traitRows, -1, drop = FALSE]
rownames(datTraits) <- Samples
```


# CHUNK 14-15: Soft threshold
```{r soft-threshold, fig.width = 10, fig.height = 5}
while (!is.null(dev.list())) dev.off()
spt <- pickSoftThreshold(expression.data, networkType = "signed hybrid", verbose = 5)
par(mfrow = c(1, 2), mar = c(4, 4, 2, 1))
plot(spt$fitIndices[, 1], -sign(spt$fitIndices[, 3]) * spt$fitIndices[, 2],
     xlab = "Soft threshold (β)", ylab = "Scale-free fit (signed R²)",
     type = "n", main = "Scale independence")
text(spt$fitIndices[, 1], -sign(spt$fitIndices[, 3]) * spt$fitIndices[, 2],
     labels = spt$fitIndices[, 1], col = "red")
abline(h = 0.85, col = "red")
plot(spt$fitIndices[, 1], spt$fitIndices[, 5],
     xlab = "Soft threshold (β)", ylab = "Mean connectivity",
     type = "n", main = "Mean connectivity")
text(spt$fitIndices[, 1], spt$fitIndices[, 5],
     labels = spt$fitIndices[, 1], col = "red")
softPower <- 6
```


# CHUNK 16-17: Adjacency + TOM
```{r adjacency-tom}
adjacency <- adjacency(expression.data, power = softPower, type = "signed hybrid")
TOM <- TOMsimilarity(adjacency, TOMType = "signed")
TOM.dissimilarity <- 1 - TOM
```

# CHUNK 18-22: Module detection + merge
```{r modules, fig.width = 12, fig.height = 7}
geneTree <- hclust(as.dist(TOM.dissimilarity), method = "average")
Modules <- cutreeDynamic(dendro = geneTree, distM = TOM.dissimilarity,
                         deepSplit = 1, pamRespectsDendro = FALSE,
                         minClusterSize = 50)
ModuleColors <- labels2colors(Modules)
MElist <- moduleEigengenes(expression.data, colors = ModuleColors)
merge        <- mergeCloseModules(expression.data, ModuleColors, cutHeight = 0.25)
mergedColors <- merge$colors
mergedMEs    <- merge$newMEs
names(mergedColors) <- colnames(expression.data)
cat("Merged modules:\n"); print(table(mergedColors))
plotDendroAndColors(geneTree, cbind(ModuleColors, mergedColors),
                    c("Original", "Merged"),
                    dendroLabels = FALSE, hang = 0.03, addGuide = TRUE,
                    guideHang = 0.05,
                    main = "Gene dendrogram — original vs merged")
```


# CHUNK 23-24: Module-trait bicor
```{r module-trait}
common_samples    <- intersect(rownames(mergedMEs), rownames(datTraits))
mergedMEs_matched <- mergedMEs[common_samples, ]
traitData         <- datTraits[common_samples, , drop = FALSE]
bicorResult <- bicorAndPvalue(mergedMEs_matched, traitData,
                              use = "pairwise.complete.obs",
                              pearsonFallback = "individual",
                              maxPOutliers = 0.1)
moduleTraitCor    <- bicorResult$bicor
moduleTraitPvalue <- bicorResult$p
moduleTraitPadj   <- matrix(
  p.adjust(as.vector(moduleTraitPvalue), method = "BH"),
  nrow = nrow(moduleTraitPvalue), ncol = ncol(moduleTraitPvalue),
  dimnames = dimnames(moduleTraitPvalue))
cat("FDR < 0.05:", sum(moduleTraitPadj < 0.05, na.rm = TRUE), "\n")
```

# CHUNK 25: Module-trait heatmap
```{r heatmap, fig.height = 9, fig.width = 11}
sig_stars <- ifelse(moduleTraitPadj < 0.001, "***",
              ifelse(moduleTraitPadj < 0.01,  "**",
              ifelse(moduleTraitPadj < 0.05,  "*",
              ifelse(moduleTraitPadj < 0.20,  "~", ""))))
textMatrix <- matrix(paste0(signif(moduleTraitCor, 2), "\n", sig_stars),
                     nrow = nrow(moduleTraitCor))
par(mar = c(6, 9, 5, 3))
labeledHeatmap(Matrix = moduleTraitCor,
               xLabels = colnames(traitData), yLabels = names(mergedMEs_matched),
               ySymbols = names(mergedMEs_matched), colorLabels = FALSE,
               colors = blueWhiteRed(50), textMatrix = textMatrix,
               setStdMargins = FALSE, cex.text = 0.45, zlim = c(-1, 1),
               main = paste0("Module-Trait Relationships (n = ",
                             length(common_samples), ")\n",
                             "*** FDR<0.001  ** FDR<0.01  * FDR<0.05  ~ FDR<0.20"))
```


# CHUNK 26: Significant associations table
```{r sig-table}
sig_idx <- which(moduleTraitPadj < 0.05, arr.ind = TRUE)
results_table <- data.frame(
  Module  = rownames(moduleTraitCor)[sig_idx[, 1]],
  Trait   = colnames(moduleTraitCor)[sig_idx[, 2]],
  bicor_r = round(moduleTraitCor[sig_idx], 3),
  p_raw   = signif(moduleTraitPvalue[sig_idx], 3),
  FDR     = signif(moduleTraitPadj[sig_idx], 3),
  Stars   = sig_stars[sig_idx])
results_table <- results_table[!grepl("grey", results_table$Module, ignore.case = TRUE), ]
results_table <- results_table[order(results_table$FDR), ]
rownames(results_table) <- NULL
print(results_table)
write.csv(results_table, file.path(OUTDIR, "module_trait_significant.csv"),
          row.names = FALSE)
```

# CHUNK 26B: Direction-aware case_control module summary
```{r}
cc_idx <- which(colnames(moduleTraitCor) == "case_control")

direction_summary <- data.frame(
  module    = gsub("^ME", "", rownames(moduleTraitCor)),
  bicor_r   = round(moduleTraitCor[, cc_idx], 3),
  p_raw     = signif(moduleTraitPvalue[, cc_idx], 3),
  FDR       = signif(moduleTraitPadj[, cc_idx], 3),
  direction = ifelse(moduleTraitCor[, cc_idx] > 0,
                     "UP in cases", "DOWN in cases"),
  stringsAsFactors = FALSE)

# Drop grey
direction_summary <- direction_summary[direction_summary$module != "grey", ]
direction_summary <- direction_summary[order(direction_summary$bicor_r), ]

cat("\n--- All modules ranked by case_control direction ---\n")
print(direction_summary)

cat("\n--- Top 5 case-DOWN modules ---\n")
print(head(direction_summary, 5))

cat("\n--- Top 5 case-UP modules ---\n")
print(tail(direction_summary, 5))

write.csv(direction_summary,
          file.path(OUTDIR, "module_case_direction_summary.csv"),
          row.names = FALSE)

# Exploratory: case-DOWN modules at relaxed FDR < 0.20
exploratory_down <- direction_summary[
  direction_summary$bicor_r < 0 & direction_summary$FDR < 0.20, ]

cat("\n--- Exploratory case-DOWN modules (FDR < 0.20) ---\n")
if (nrow(exploratory_down) > 0) {
  print(exploratory_down)
} else {
  cat("None found.\n")
}

# Save for Discussion section
write.csv(exploratory_down,
          file.path(DIR_ENRICH, "case_down_exploratory_FDR020.csv"),
          row.names = FALSE)
```

# CHUNK 27: Biotype composition  (margins fixed for inline rendering)
```{r biotype-composition, fig.height = 7, fig.width = 14}
biotype_lookup <- setNames(anno$gene_biotype, anno$ensembl_gene_id)
gene_biotypes  <- biotype_lookup[colnames(expression.data)]
biotype_grouped <- dplyr::case_when(
  gene_biotypes == "lncRNA"         ~ "lncRNA",
  gene_biotypes == "protein_coding" ~ "protein_coding",
  is.na(gene_biotypes)              ~ NA_character_,
  TRUE                              ~ "other_ncRNA")
module_biotype_df <- data.frame(gene_id = names(mergedColors),
                                module  = mergedColors,
                                biotype = biotype_grouped) %>%
  dplyr::filter(!is.na(biotype))
module_counts <- as.data.frame(
  table(module = module_biotype_df$module, biotype = module_biotype_df$biotype))
module_wide <- reshape(module_counts, idvar = "module",
                       timevar = "biotype", direction = "wide")
colnames(module_wide) <- gsub("Freq\\.", "", colnames(module_wide))
module_wide[is.na(module_wide)] <- 0
if (!"other_ncRNA" %in% colnames(module_wide)) module_wide$other_ncRNA <- 0
module_wide$mapped_total <- rowSums(module_wide[, c("lncRNA", "protein_coding")])
module_wide$pct_lncRNA   <- round(100 * module_wide$lncRNA / module_wide$mapped_total, 1)
module_wide$pct_coding   <- round(100 * module_wide$protein_coding / module_wide$mapped_total, 1)
module_wide <- module_wide[order(-module_wide$pct_lncRNA), ]
write.csv(module_wide, file.path(OUTDIR, "module_biotype_composition.csv"),
          row.names = FALSE)

plot_mat <- t(as.matrix(
  module_wide[, c("lncRNA", "protein_coding")] / module_wide$mapped_total * 100))
colnames(plot_mat) <- module_wide$module
par(mar = c(8, 4, 3, 2), xpd = TRUE)            # fixed margins
barplot(plot_mat, col = c("#E05C5C", "#5B8DB8"), border = NA, las = 2,
        cex.names = 0.55, ylab = "% of mapped module genes", ylim = c(0, 100),
        main = "Biotype Composition per WGCNA Module (sorted by lncRNA %)")
abline(h = 50, lty = 2, col = "black", lwd = 1.2)
legend("right", inset = c(-0.05, 0), legend = c("lncRNA", "protein_coding"),
       fill = c("#E05C5C", "#5B8DB8"), border = NA, bty = "n")
```


# CHUNK 28: Module Membership + Gene Significance
```{r gs-mm}
modNames <- substring(names(mergedMEs), 3)
MM_result <- bicorAndPvalue(expression.data, mergedMEs, use = "p")
geneModuleMembership <- as.data.frame(MM_result$bicor)
MMpvalue             <- as.data.frame(MM_result$p)
names(geneModuleMembership) <- paste0("MM.",   modNames)
names(MMpvalue)             <- paste0("p.MM.", modNames)
GS_result <- bicorAndPvalue(expression.data, traitData, use = "p")
geneTraitSignificance <- as.data.frame(GS_result$bicor)
GSvalue               <- as.data.frame(GS_result$p)
names(geneTraitSignificance) <- paste0("GS.",   names(traitData))
names(GSvalue)               <- paste0("p.GS.", names(traitData))
```


# CHUNK 29: MM vs GS scatter (auto)
```{r mm-gs-auto, fig.width = 6, fig.height = 5}
if (nrow(results_table) > 0) {
  sig_pairs <- results_table[, c("Module", "Trait")]
  sig_pairs$Module <- gsub("^ME", "", sig_pairs$Module)
  sig_pairs <- sig_pairs[sig_pairs$Module %in% modNames, ]
  sig_pairs <- sig_pairs[paste0("GS.", sig_pairs$Trait) %in%
                         colnames(geneTraitSignificance), ]
  for (i in seq_len(nrow(sig_pairs))) {
    mod    <- sig_pairs$Module[i]
    trait  <- sig_pairs$Trait[i]
    column <- match(mod, modNames)
    moduleGenes <- mergedColors == mod
    if (is.na(column) || sum(moduleGenes) == 0) next
    verboseScatterplot(
      abs(geneModuleMembership[moduleGenes, column]),
      abs(geneTraitSignificance[moduleGenes, paste0("GS.", trait)]),
      xlab = paste("MM in", mod), ylab = paste("GS for", trait),
      main = paste0("MM vs GS • ", mod, " × ", trait), col = mod)
  }
}
```


# CHUNK 30: geneInfo table
```{r gene-info}
probes       <- colnames(expression.data)
probes2annot <- match(probes, anno$ensembl_gene_id)
geneInfo0 <- data.frame(
  ensembl_id   = probes,
  geneSymbol   = anno$hgnc_symbol[probes2annot],
  gene_biotype = anno$gene_biotype[probes2annot],
  moduleColor  = mergedColors,
  geneTraitSignificance,
  GSvalue,
  stringsAsFactors = FALSE)
modOrder <- order(-abs(cor(mergedMEs, traitData$case_control, use = "p")))
for (k in seq_len(ncol(geneModuleMembership))) {
  oldNames  <- names(geneInfo0)
  geneInfo0 <- data.frame(geneInfo0,
                          geneModuleMembership[, modOrder[k]],
                          MMpvalue[, modOrder[k]])
  names(geneInfo0) <- c(oldNames,
                        paste0("MM.",   modNames[modOrder[k]]),
                        paste0("p.MM.", modNames[modOrder[k]]))
}
geneOrder <- order(geneInfo0$moduleColor, -abs(geneInfo0$GS.case_control))
geneInfo  <- geneInfo0[geneOrder, ]
write.csv(geneInfo, file.path(OUTDIR, "geneInfo_full.csv"), row.names = FALSE)
cat("geneInfo saved:", nrow(geneInfo), "rows ×", ncol(geneInfo), "columns\n")
```


# CHUNK 31: Top hub genes per significant module
```{r hub-table}
sig_modules <- unique(gsub("^ME", "", results_table$Module))
sig_modules <- setdiff(sig_modules, "grey")
for (mod in sig_modules) {
  mm_col <- paste0("MM.", mod)
  if (!mm_col %in% colnames(geneInfo)) next
  cat("\n-- Module:", mod, "--\n")
  sub <- geneInfo[geneInfo$moduleColor == mod, ]
  sub <- sub[order(-abs(sub[[mm_col]])), ]
  print(head(sub[, c("geneSymbol", "gene_biotype", mm_col,
                     "GS.case_control", "GS.ER", "GS.grade")], 5))
}
```


# CHUNK 32: Consolidated fgsea — single computation across 4 collections
```{r fgsea-master}
all_modules <- setdiff(unique(geneInfo$moduleColor), "grey")
cat("Non-grey modules to test:", length(all_modules), "\n")

load_collection <- function(coll, sub = NULL) {
  args <- list(species = "Homo sapiens", collection = coll)
  if (!is.null(sub)) args$subcollection <- sub
  sets <- do.call(msigdbr, args)
  split(sets$ensembl_gene, sets$gs_name)
}
collections <- list(
  Hallmark      = load_collection("H"),
  GO_BP         = load_collection("C5", "GO:BP"),
  Reactome      = load_collection("C2", "CP:REACTOME"),
  C1_positional = load_collection("C1"))
for (n in names(collections))
  cat(sprintf("  %-15s: %d gene sets\n", n, length(collections[[n]])))

# Helper — runs fgsea on one module against one collection
# scoreType = "pos" silences within-module MM-ranking warnings (all values > 0)
run_fgsea_module <- function(mod, paths) {
  mm_col <- paste0("MM.", mod)
  if (!mm_col %in% colnames(geneInfo)) return(NULL)
  mod_df <- geneInfo[geneInfo$moduleColor == mod, ]
  mr <- setNames(mod_df[[mm_col]], mod_df$ensembl_id)
  mr <- sort(mr[!is.na(mr)], decreasing = TRUE)
  if (length(mr) < 10) return(NULL)
  res <- suppressWarnings(tryCatch(
    fgsea(pathways = paths, stats = mr,
          minSize = 5, maxSize = 500, eps = 0,
          scoreType = "std"),
    error = function(e) NULL))
  if (is.null(res) || nrow(res) == 0) return(NULL)
  res$module <- mod
  res
}

# Run all 4 × 38 combinations once and cache
fgsea_all <- list()
set.seed(42)
for (coll_name in names(collections)) {
  cat("\nRunning", coll_name, "...\n")
  fgsea_all[[coll_name]] <- lapply(all_modules, run_fgsea_module,
                                   paths = collections[[coll_name]])
  names(fgsea_all[[coll_name]]) <- all_modules
}
saveRDS(fgsea_all, file.path(INTERMEDIATE, "fgsea_results_master.rds"))

# Build long-format data frame with leadingEdge flattened to avoid CSV errors
long_df <- do.call(rbind, lapply(names(fgsea_all), function(coll) {
  do.call(rbind, lapply(all_modules, function(mod) {
    r <- fgsea_all[[coll]][[mod]]
    if (is.null(r) || nrow(r) == 0) return(NULL)
    d <- as.data.frame(r)
    d$collection  <- coll
    d$leadingEdge <- sapply(d$leadingEdge, function(x) paste(x, collapse = ";"))
    d
  }))
}))
write.csv(long_df, file.path(DIR_ENRICH, "fgsea_all_results.csv"), row.names = FALSE)
cat("\nAll fgsea results saved:", nrow(long_df), "rows\n")
```
# CHUNK 32a: Hallmark enrichment bubble heatmap across all modules
```{r hallmark-bubble, fig.width = 11, fig.height = 9}
hm <- subset(long_df, collection == "Hallmark")
hm$pathway <- gsub("^HALLMARK_", "", hm$pathway)
keep <- unique(hm$pathway[hm$padj < 0.05])          # pathways significant in >=1 module
hm <- hm[hm$pathway %in% keep, ]
p_bubble <- ggplot(hm, aes(x = module, y = pathway, size = -log10(padj), colour = NES)) +
  geom_point(alpha = 0.9) +
  scale_colour_gradient2(low = "blue", mid = "grey90", high = "red", midpoint = 0) +
  scale_size_continuous(range = c(0.5, 6), name = expression(-log[10]~padj)) +
  labs(x = NULL, y = NULL, title = "MSigDB Hallmark enrichment across WGCNA modules") +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 7),
        axis.text.y = element_text(size = 7))
ggsave(file.path(DIR_ENRICH, "hallmark_bubble_heatmap.png"), p_bubble,
       width = 11, height = 9, dpi = 300)
```

# CHUNK 32b: Enrichment dotplot for the three recurrence-associated modules
```{r recurrence-dotplot, fig.width = 9, fig.height = 8}
recurrence_modules <- c("paleturquoise", "violet", "darkorange")
enr <- subset(long_df, module %in% recurrence_modules & padj < 0.05)
enr$term <- gsub("^(HALLMARK|GOBP|REACTOME)_", "", enr$pathway)
enr <- enr[order(enr$module, enr$padj), ]
top_terms <- do.call(rbind, by(enr, enr$module, function(d) head(d, 8)))
top_terms$module <- factor(top_terms$module, levels = recurrence_modules)
top_terms$term   <- factor(top_terms$term, levels = unique(top_terms$term[order(top_terms$NES)]))
p_dot <- ggplot(top_terms, aes(x = module, y = term, size = size, colour = -log10(padj))) +
  geom_point() +
  scale_colour_viridis_c(option = "C", name = expression(-log[10]~padj)) +
  scale_size_continuous(range = c(1, 7), name = "gene-set size") +
  labs(x = NULL, y = NULL, title = "Enrichment in recurrence-associated modules") +
  theme_bw() + theme(axis.text.y = element_text(size = 7))
ggsave(file.path(DIR_ENRICH, "enrichment_dotplot_CCmodules.png"), p_dot,
       width = 9, height = 8, dpi = 300)
```

# CHUNK 32c: GSEA running-enrichment ("mountain") plots for the recurrence modules
```{r gsea-mountain, fig.width = 6, fig.height = 4}
recurrence_modules <- c("darkorange", "paleturquoise", "violet")
for (mod in recurrence_modules) {
  mm_col <- paste0("MM.", mod)
  md <- geneInfo[geneInfo$moduleColor == mod, ]
  mr <- sort(setNames(md[[mm_col]], md$ensembl_id)[!is.na(md[[mm_col]])], decreasing = TRUE)
  top <- subset(long_df, collection == "Hallmark" & module == mod)
  top <- top[order(top$padj, top$pval), ][1, ]
  p <- fgsea::plotEnrichment(collections$Hallmark[[top$pathway]], mr) +
       ggplot2::labs(title = paste0(mod, " — ", top$pathway),
                     subtitle = sprintf("NES = %.2f, padj = %.2g", top$NES, top$padj))
  ggsave(file.path(DIR_ENRICH, paste0("gsea_mountain_", mod, ".png")), p,
         width = 6, height = 4, dpi = 300)
}
```


# CHUNK 33: Master annotation table
```{r fgsea-rescue-summary}
hit_counts <- long_df %>%
  dplyr::group_by(module, collection) %>%
  dplyr::summarise(n_sig    = sum(padj < 0.05, na.rm = TRUE),
                   n_tested = dplyr::n(), .groups = "drop")
hit_wide <- hit_counts %>%
  dplyr::select(module, collection, n_sig) %>%
  tidyr::pivot_wider(names_from = collection, values_from = n_sig, values_fill = 0)

mod_size <- geneInfo %>%
  dplyr::filter(moduleColor != "grey") %>%
  dplyr::group_by(moduleColor) %>%
  dplyr::summarise(n_genes = dplyr::n(),
                   pct_lncRNA = round(100 * sum(gene_biotype == "lncRNA") /
                                        dplyr::n(), 1),
                   .groups = "drop") %>%
  dplyr::rename(module = moduleColor)

trait_sig <- results_table %>%
  dplyr::mutate(module = gsub("^ME", "", Module)) %>%
  dplyr::group_by(module) %>%
  dplyr::summarise(n_traits_sig = dplyr::n_distinct(Trait),
                   significant_traits = paste(unique(Trait), collapse = ", "),
                   .groups = "drop")

master <- mod_size %>%
  dplyr::left_join(trait_sig, by = "module") %>%
  dplyr::left_join(hit_wide,  by = "module") %>%
  dplyr::mutate(across(c(n_traits_sig, Hallmark, GO_BP, Reactome, C1_positional),
                       ~ tidyr::replace_na(.x, 0)),
                significant_traits = tidyr::replace_na(significant_traits, "—"))

master <- master %>%
  dplyr::mutate(rescued_by = dplyr::case_when(
    Hallmark > 0      ~ "Hallmark",
    GO_BP > 0         ~ "GO BP",
    Reactome > 0      ~ "Reactome",
    C1_positional > 0 ~ "C1 positional",
    TRUE              ~ "None")) %>%
  dplyr::mutate(scope = dplyr::case_when(
    n_traits_sig > 0 & rescued_by != "None" ~ "Include: trait + biology",
    n_traits_sig > 0                        ~ "Include: trait only",
    rescued_by != "None"                    ~ "Include: biology only",
    TRUE                                    ~ "Exclude: no trait, no biology")) %>%
  dplyr::arrange(scope, dplyr::desc(n_genes))

print(master)
write.csv(master, file.path(OUTDIR, "master_annotation_table.csv"), row.names = FALSE)
cat("\nScope counts:\n"); print(table(master$scope))
```


# CHUNK 34: Master annotation figure
```{r master-figure, fig.height = 10, fig.width = 12}
master_plot <- master
master_plot$module     <- factor(master_plot$module,
                                 levels = master_plot$module[order(master_plot$n_genes)])
master_plot$rescued_by <- factor(master_plot$rescued_by,
                                 levels = c("Hallmark","GO BP","Reactome",
                                            "C1 positional","None"))
p_master <- ggplot(master_plot, aes(x = module, y = n_genes, fill = rescued_by)) +
  geom_col() +
  geom_text(aes(label = ifelse(n_traits_sig > 0,
                               paste0("★", n_traits_sig), "")),
            hjust = -0.2, size = 3) +
  scale_fill_manual(values = c("Hallmark"      = "#d73027",
                               "GO BP"         = "#fc8d59",
                               "Reactome"      = "#fdae61",
                               "C1 positional" = "#fee090",
                               "None"          = "grey75")) +
  coord_flip() +
  labs(title    = "Module annotation status across enrichment strategies",
       subtitle = "Bar = module size; colour = first database providing significance; ★n = significant traits",
       x = NULL, y = "Number of genes", fill = "Rescued by") +
  theme_bw(base_size = 10) +
  theme(legend.position = "bottom",
        plot.title = element_text(face = "bold"))
print(p_master)
ggsave(file.path(OUTDIR, "master_annotation_status.png"),
       p_master, width = 12, height = 10, dpi = 200, bg = "white")
```


# CHUNK 35: All-modules Hallmark heatmap
```{r hallmark-heatmap-all, fig.height = 12, fig.width = 14}
hm <- long_df %>%
  dplyr::filter(collection == "Hallmark") %>%
  dplyr::mutate(pathway = gsub("^HALLMARK_", "", pathway))
p_hm <- ggplot(hm, aes(x = module, y = pathway, fill = NES)) +
  geom_tile(color = "grey90") +
  scale_fill_gradient2(low = "#4575b4", mid = "white", high = "#d73027",
                       midpoint = 0, limits = c(-3, 3),
                       oob = scales::squish, name = "NES") +
  geom_text(aes(label = ifelse(padj < 0.05, "*", "")), size = 3, vjust = 0.7) +
  theme_minimal(base_size = 9) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  labs(title    = "Hallmark enrichment across all modules",
       subtitle = "* = padj < 0.05",
       x = NULL, y = NULL)
print(p_hm)
ggsave(file.path(OUTDIR, "hallmark_heatmap_all_modules.png"),
       p_hm, width = 14, height = 12, dpi = 200, bg = "white")
```


# CHUNK 36: Per-module Hallmark barplots (significant only)
```{r hallmark-barplots, fig.width = 8, fig.height = 5}
hallmark_sig_mods <- master$module[master$Hallmark > 0]
for (mod in as.character(hallmark_sig_mods)) {
  d <- long_df %>%
    dplyr::filter(collection == "Hallmark", module == mod, padj < 0.05)
  if (nrow(d) == 0) next
  d$pathway   <- gsub("^HALLMARK_", "", d$pathway)
  d$direction <- ifelse(d$NES > 0, "Activated", "Suppressed")
  p <- ggplot(d, aes(reorder(pathway, NES), NES, fill = direction)) +
    geom_col() +
    scale_fill_manual(values = c("Activated" = "#d73027", "Suppressed" = "#4575b4")) +
    coord_flip() +
    labs(title = paste("Hallmark —", mod), x = NULL, y = "NES", fill = NULL) +
    theme_bw(base_size = 11) + theme(legend.position = "bottom")
  print(p)
}
```

# CHUNK 36B Plots top 10 Hallmark pathways for every module that did NOT have
# any padj < 0.05 hits — showing NES "for reference" so absence
```{r hallmark-barplots-below-threshold, fig.width = 8, fig.height = 5}
nonsig_hallmark_mods <- master$module[master$Hallmark == 0]

for (mod in as.character(nonsig_hallmark_mods)) {
  d <- long_df %>%
    dplyr::filter(collection == "Hallmark", module == mod)
  if (nrow(d) == 0) next
  
  # Top 10 by |NES| (no padj filter — by definition none pass)
  d <- d %>%
    dplyr::arrange(dplyr::desc(abs(NES))) %>%
    dplyr::slice_head(n = 10)
  d$pathway   <- gsub("^HALLMARK_", "", d$pathway)
  d$direction <- ifelse(d$NES > 0, "Activated", "Suppressed")

  p <- ggplot(d, aes(reorder(pathway, NES), NES, fill = direction)) +
    geom_col(alpha = 0.7) +
    scale_fill_manual(values = c("Activated"  = "#d73027",
                                 "Suppressed" = "#4575b4")) +
    coord_flip() +
    labs(title    = paste("Hallmark —", mod, "(below threshold)"),
         subtitle = "Top 10 by |NES|; no pathway reached padj < 0.05",
         x = NULL, y = "Normalised Enrichment Score (NES)", fill = NULL) +
    theme_bw(base_size = 11) +
    theme(legend.position = "bottom",
          plot.subtitle   = element_text(face = "italic", colour = "grey40"))
  print(p)
}

cat("\nPlotted", length(nonsig_hallmark_mods),
    "below-threshold modules.\n")
```

# CHUNK 37: Rescue grid for GO BP / Reactome / C1 modules
```{r rescue-plots, fig.width = 8, fig.height = 5}
rescued_mods <- master %>%
  dplyr::filter(rescued_by %in% c("GO BP","Reactome","C1 positional")) %>%
  dplyr::arrange(rescued_by, module)

for (i in seq_len(nrow(rescued_mods))) {
  mod       <- as.character(rescued_mods$module[i])
  best      <- as.character(rescued_mods$rescued_by[i])
  best_coll <- gsub(" ", "_", best)

  d <- long_df %>%
    dplyr::filter(module == mod, collection == best_coll, padj < 0.05) %>%
    dplyr::arrange(dplyr::desc(abs(NES))) %>%
    dplyr::slice_head(n = 15)
  if (nrow(d) == 0) next

  d$pathway   <- tolower(gsub("_", " ",
                              gsub("^(HALLMARK_|GOBP_|REACTOME_)", "", d$pathway)))
  d$direction <- ifelse(d$NES > 0, "Activated", "Suppressed")

  p <- ggplot(d, aes(reorder(pathway, NES), NES, fill = direction)) +
    geom_col(alpha = 0.85) +
    scale_fill_manual(values = c("Activated"  = "#d73027",
                                 "Suppressed" = "#4575b4")) +
    coord_flip() +
    labs(title    = paste0(mod, " — ", best),
         subtitle = sprintf("Top %d pathways with padj < 0.05", nrow(d)),
         x = NULL, y = "Normalised Enrichment Score (NES)", fill = NULL) +
    theme_bw(base_size = 11) +
    theme(legend.position = "bottom",
          plot.title    = element_text(face = "bold"),
          plot.subtitle = element_text(face = "italic", colour = "grey40"))
  print(p)
}

cat("\nPlotted", nrow(rescued_mods), "rescued modules individually.\n")
```

# CHUNK 38: Cytoscape network export
```{r cytoscape-export}
mergedColors <- setNames(as.character(mergedColors), colnames(expression.data))
sig_mods <- intersect(sig_modules, unique(mergedColors))
sig_mods <- setdiff(sig_mods, "grey")
probes       <- colnames(expression.data)
probe_symbol <- anno$hgnc_symbol[match(probes, anno$ensembl_gene_id)]
probe_symbol[is.na(probe_symbol) | probe_symbol == ""] <-
  probes[is.na(probe_symbol) | probe_symbol == ""]
nTop <- 30
hub_summary <- data.frame()
for (mod in sig_mods) {
  inModule   <- mergedColors == mod
  modProbes  <- probes[inModule]
  modSymbols <- probe_symbol[inModule]
  if (sum(inModule) == 0) next
  modTOM <- TOM[inModule, inModule]
  rownames(modTOM) <- colnames(modTOM) <- modSymbols
  IMConn <- rowSums(modTOM) - 1
  nSelect <- min(nTop, sum(inModule))
  topIdx <- rank(-IMConn) <= nSelect
  ord    <- order(-IMConn[topIdx])
  exportNetworkToCytoscape(
    modTOM[topIdx, topIdx],
    edgeFile = file.path(OUTDIR, paste0("Cytoscape_edges_", mod, ".txt")),
    nodeFile = file.path(OUTDIR, paste0("Cytoscape_nodes_", mod, ".txt")),
    weighted = TRUE, threshold = 0.02,
    nodeNames = modSymbols[topIdx][ord],
    nodeAttr  = rep(mod, nSelect))
  hub_summary <- rbind(hub_summary, data.frame(
    module        = mod,
    rank          = seq_len(nSelect),
    gene_id       = modProbes[topIdx][ord],
    gene_symbol   = modSymbols[topIdx][ord],
    intramod_conn = round(IMConn[topIdx][ord], 4)))
}
write.csv(hub_summary, file.path(OUTDIR, "hub_summary.csv"), row.names = FALSE)
cat("Cytoscape files exported for", length(sig_mods), "modules\n")
```

# CHUNK 39: DEG × WGCNA integration
```{r deg-integration}
DEG_PATH <- file.path(DIR_CCDEG, "allsamples_DEgenes.tsv")
deg <- read.table(DEG_PATH, header = TRUE, sep = "\t",
                  row.names = 1, stringsAsFactors = FALSE)
common_ids <- intersect(rownames(deg), rownames(geneInfo))
deg_mapped <- cbind(geneInfo[common_ids, ],
                    deg[common_ids, c("log2FoldChange", "padj", "diffexpressed")])
deg_mapped$own_MM <- sapply(seq_len(nrow(deg_mapped)), function(i) {
  col <- paste0("MM.", deg_mapped$moduleColor[i])
  if (col %in% colnames(deg_mapped)) deg_mapped[i, col] else NA_real_
})
print(table(deg_mapped$moduleColor, deg_mapped$diffexpressed))

enrich_df <- do.call(rbind, lapply(
  setdiff(unique(geneInfo$moduleColor), "grey"),
  function(mod) {
    n_mod <- sum(geneInfo$moduleColor == mod)
    n_oth <- nrow(geneInfo) - n_mod
    n_up <- sum(deg_mapped$moduleColor == mod & deg_mapped$diffexpressed == "UP",   na.rm = TRUE)
    n_dn <- sum(deg_mapped$moduleColor == mod & deg_mapped$diffexpressed == "DOWN", na.rm = TRUE)
    if (n_up + n_dn == 0) return(NULL)
    total_up <- sum(deg_mapped$diffexpressed == "UP",   na.rm = TRUE)
    total_dn <- sum(deg_mapped$diffexpressed == "DOWN", na.rm = TRUE)
    p_up <- tryCatch(fisher.test(matrix(c(n_up, total_up - n_up,
                                          n_mod - n_up, n_oth - (total_up - n_up)),
                                        2, 2), alternative = "greater")$p.value,
                     error = function(e) NA)
    p_dn <- tryCatch(fisher.test(matrix(c(n_dn, total_dn - n_dn,
                                          n_mod - n_dn, n_oth - (total_dn - n_dn)),
                                        2, 2), alternative = "greater")$p.value,
                     error = function(e) NA)
    data.frame(module = mod, n_genes = n_mod, n_UP = n_up, n_DOWN = n_dn,
               p_UP = p_up, p_DOWN = p_dn)
  }))
enrich_df$FDR_UP   <- p.adjust(enrich_df$p_UP,   method = "BH")
enrich_df$FDR_DOWN <- p.adjust(enrich_df$p_DOWN, method = "BH")
write.csv(enrich_df, file.path(DIR_ENRICH, "deg_module_enrichment.csv"), row.names = FALSE)
```


# CHUNK 40: MM vs GS scatter coloured by DEG
```{r deg-scatter-auto, fig.width = 6, fig.height = 5}
plot_MM_vs_GS <- function(mod, trait) {
  mm_col <- paste0("MM.", mod); gs_col <- paste0("GS.", trait)
  if (!all(c(mm_col, gs_col) %in% colnames(geneInfo))) return(invisible(NULL))
  pd <- data.frame(
    MM = geneInfo[[mm_col]], GS = geneInfo[[gs_col]],
    sym = geneInfo$geneSymbol,
    dir = ifelse(rownames(geneInfo) %in%
                 rownames(deg_mapped[deg_mapped$diffexpressed == "UP", ]),   "UP",
          ifelse(rownames(geneInfo) %in%
                 rownames(deg_mapped[deg_mapped$diffexpressed == "DOWN", ]), "DOWN", "none")))
  pd <- pd[geneInfo$moduleColor == mod & !is.na(pd$MM) & !is.na(pd$GS), ]
  if (nrow(pd) == 0) return(invisible(NULL))
  pd$label <- ifelse(pd$dir != "none", pd$sym,
                     ifelse(rank(-abs(pd$MM)) <= 5, pd$sym, ""))
  r <- round(cor(pd$MM, pd$GS, method = "spearman"), 3)
  print(
    ggplot(pd, aes(MM, GS)) +
      geom_point(data = subset(pd, dir == "none"),
                 colour = "grey70", size = 1, alpha = 0.4) +
      geom_point(data = subset(pd, dir == "UP"),
                 colour = "#C00000", size = 2.5, shape = 17) +
      geom_point(data = subset(pd, dir == "DOWN"),
                 colour = "#0070C0", size = 2.5, shape = 25) +
      geom_smooth(method = "lm", se = TRUE,
                  colour = "black", linewidth = 0.7, linetype = "dashed") +
      geom_text_repel(aes(label = label), size = 3, max.overlaps = 20,
                      colour = ifelse(pd$dir == "UP",   "#C00000",
                               ifelse(pd$dir == "DOWN", "#0070C0", "grey40"))) +
      labs(title    = paste0("ME", mod, " · MM vs GS.", trait),
           subtitle = paste0("n=", nrow(pd),
                             " | DEGs=", sum(pd$dir != "none"),
                             " | Spearman r=", r),
           x = mm_col, y = gs_col,
           caption = "▲ UP-DEG  ▼ DOWN-DEG  ● background") +
      theme_bw(base_size = 12) +
      theme(plot.title = element_text(face = "bold")))
}
sig_pairs_scatter <- results_table[, c("Module", "Trait")]
sig_pairs_scatter$Module <- gsub("^ME", "", sig_pairs_scatter$Module)
sig_pairs_scatter <- sig_pairs_scatter[
  sig_pairs_scatter$Module %in% modNames &
  paste0("GS.", sig_pairs_scatter$Trait) %in% colnames(geneInfo), ]
for (i in seq_len(nrow(sig_pairs_scatter))) {
  plot_MM_vs_GS(sig_pairs_scatter$Module[i],
                sig_pairs_scatter$Trait[i])
}
```


# CHUNK 41: DEG volcano coloured by WGCNA module
```{r volcano-by-module, fig.width = 9, fig.height = 7}
deg_mapped %>%
  mutate(nlp = -log10(padj),
         label = case_when(
           abs(own_MM) >= 0.75 & padj < 0.01           ~ geneSymbol,
           -log10(padj) > 5                             ~ geneSymbol,
           gene_biotype == "protein_coding" &
             abs(log2FoldChange) > 1.5                 ~ geneSymbol,
           TRUE                                         ~ "")) %>%
  ggplot(aes(log2FoldChange, nlp, colour = moduleColor)) +
  geom_point(size = 2.5, alpha = 0.85) +
  geom_hline(yintercept = -log10(0.05), linetype = "dashed", colour = "grey50") +
  geom_vline(xintercept = c(-0.5, 0.5), linetype = "dashed", colour = "grey50") +
  geom_text_repel(aes(label = label), size = 3, max.overlaps = 25,
                  show.legend = FALSE) +
  scale_colour_identity() +
  labs(title = "DEG Volcano coloured by WGCNA module",
       x = "log2FC ('left' control vs 'right' cases)", y = "-log10(padj)") +
  theme_bw(base_size = 12) +
  theme(plot.title = element_text(face = "bold"))
```


# CHUNK 42: Save workspace
```{r save-workspace}
save(expression.data, mergedColors, mergedMEs, TOM,
     geneInfo, results_table, sig_modules,
     fgsea_all, long_df, master,
     deg_mapped, enrich_df, hub_summary,
     file = file.path(INTERMEDIATE, "WGCNA_v5_workspace.RData"))
cat("Workspace saved to", file.path(INTERMEDIATE, "WGCNA_v5_workspace.RData"), "\n")
cat("\n=== Files written to OUTDIR ===\n")
print(list.files(OUTDIR))
```
